% My Encryption Algorithim Article
% Henry J Schmale
% March 3, 2015

\documentclass[12pt]{article}

% Margins
\usepackage[letterpaper]{geometry}
\geometry{
  top    = 1.0in,
  bottom = 1.0in,
  left   = 1.0in,
  right  = 1.0in}

% Font
\usepackage{times}

% Spacing
\setlength{\textheight}{9.5in}
\usepackage{setspace}
\doublespacing

% Math Formulas
\usepackage{mathtools}

% Algorithims and code Formatting
\usepackage[]{algpseudocode}
\usepackage{listings}
\usepackage{color}

\lstset{language=C++,
        basicstyle=\small,
        numbers=left,
        numberstyle=\tiny,
        stepnumber=1}

% Start of the document
\begin{document}

% TITLE
\title{Hiding an Encrypted Message Inside an Image using
      Standard File Formats}
\author{Henry J Schmale}
\date{March 2015}
\maketitle

% abstract
\begin{abstract}
  The design and implementation of an algorithim to encrypt a message
  and hide it inside of a standard image file. This paper describes
  the implementation of a simple xor rotate cipher, and then gives
  a description on how to encode the output data inside of an image.
  Additionally, this paper includes details of extracting the
  encrypted message from an image, and undoing the enryption on that
  image.
\end{abstract}

\newpage
\tableofcontents                       % Table of Contents
\listoffigures                         % Figures List
\listoftables                          % Tables list
\newpage

% MAIN CONTENT
\section{Introduction}
  Encryption is a powerful tool to protect your information, but standard
  encryption can be insecure if found. Additionally, it can be 
  incriminating if found, so that is why hiding the encrypted 
  information is more important then ever. The art and science of hiding
  information in plain sight is called stenography. This document 
  describes the design and implementation of such a system to encrypt
  and hide information inside of a standard image file.

  \subsection{A Note About the Code in this Article}
    The code samples featured in this article are written in C-style 
    C++11. This means that the code featured in this article makes use 
    of c-style arrays and direct bit manipulation. The main reason the 
    code samples featured in this article are written in C++11 is the 
    use of standardized fixed-width integers included in the standard 
    library.

\section{Encryption and Decryption}
  The encryption used in this system is a simple XOR rotation cipher.
  The cipher uses an n-byte key to encrypt and decrypt the message.
  Along with the use of a rotating key, this a relatively fast and
  secure algorithim, though probably nowhere near as secure as any of
  public-private key algorithims. However, this algorithim was not
  designed with that in mind, this algorithim was designed with the
  thought of protecting a file against a cursory overview.
  
  As this algorithim uses a rotating key to encrypt mesages there
  must be regular pattern to how the key changes. As such the
  change to the key is a simple bit rotation after each use. 

  % Equation describing the value of K to use in this iteration
  \begin{equation}
    \label{eq:keyDerive}
    K_i = rotateRight(
        \underbrace{K_{i \bmod length(K)}}_\text{Value to rotate},
        \underbrace{\frac{i}{length(K)} \bmod (sizeof(type(K)) 
            \times B)}_\text{places}) 
  \end{equation}

  Equation \eqref{eq:keyDerive} describes how to calculate the key 
  value to use in iteration \( i \) of the encryption algorithim. 
  It requires the use of a bitwise rotate operation so that no data 
  is lost, through the replacement by zeros when using standard 
  bitwise shifts. And as zeros are introduced into the key, the 
  resulting value of an XOR would no longer work as a key, because
  the encoded values would remain the same.

  This algorithim can use an n-bit key as the new key value is
  determined independtly of the length of the key and provides
  facilities for working with various key lengths. As such it can use
  a n-bit key for encryption and decryption.

  The encryption algorithim requires the use of the bitwise XOR and
  takes advantage of the speed yielded with an XOR operation, due to
  no carry bits being required, and dedicated circuitry existing on most
  modern processors. Below is a description of the algorithim in detail.

  \subsection{Encrypting a Message}
    This section describes the process of encrypting a message.
    The general process of encrypting a message using an xor rotate
    cipher is quite simple, but it becomes more complex with the
    addition of a rotating key. The general process of encryption
    using an XOR rotation cipher is descriped below. 
    
    \begin{figure}[h]
    \begin{algorithmic}[1]
    \For{$i=0; i < length(OriginalMessage); i++$}
      \State $EncryptedMesage_i =
             OriginalMessage_i \oplus Key_{i \bmod length(Key)}$
      \State $EncryptedMessage_i = rotateRight(EncryptedMessage_i)$
    \EndFor 
    \end{algorithmic}
    \caption{A Basic XOR rotation cipher}
    \end{figure}

    If there was nothing more than an XOR opperation in this algorithim,
    it would be incredibly easy to use a brute-force attack against.
    However, since a bit rotation was added into the algorithim, there
    are now \(CHAR\_BITS^2\) possible solutions to the message, making
    brute-forcing it, less of an option. However, the key is still
    vurnable to being brute-forced, so modification of the key is also
    required. Yielding, something more like the following.
    
    \begin{figure}[h]
    \begin{algorithmic}[1]
    \label{fig:ImprovRotCiph}
    \For{ $i=0; i < length(OriginalMessage); i++$}
      \State $EncryptedMessage_i =
             OriginalMessage_i \oplus Key_{i \bmod length(Key)}$
      \State $EncryptedMessage_i = rotateRight(EncryptedMessage_i)$
      \If{$( i \bmod length(Key) ) \equiv 0$}
        \For{ $j=0;j < length(Key); j++$ }
          \State $Key_j = rotateRight(Key_j)$
        \EndFor
      \EndIf
    \EndFor
    \end{algorithmic}
    \caption{An Improved XOR rotation cipher}
    \end{figure}
    
    As you can see the second example is much more confusing, as it 
    involves manipulation of the actual key. As more confusion is 
    better in cryptography the second algorithim is much better.
    Additionally, the encryption process can be described as a
    mathmatical formula, as seen in equation \eqref{eq:Encrypt}. 
    With \(K_i\) as defined in equation \eqref{eq:keyDerive}.
    
    \begin{equation}
      \label{eq:Encrypt}
      E_{i} = rotateRight(M_{i} \oplus K_{i})   
    \end{equation}

    \subsubsection{Programming an Encryption Algorithim}
    Now to implement such an algorithim in C++, we need to perform several
    checks before we can even begin. First of all, nulls must be
    checked for, as three parameters are pointers, and can be null.
    As writing to a null pointer would cause a segmentation fault, 
    it must be checked for. Therefore, the encryption function should 
    start with a couple of standard asserts or macro that does the same
    thing. Additionally, memory allocation sizes must be checked,
    otherwise the encryption function might write to
    memory that the program hasn't  been allocated, causing another
    segmentation fault. Finally, the encryption key must be copied into
    local memory as the Key is passed by pointer, and can be modified
    within that function and returned via pointer, the encryption key
    might be used in other parts of the program and should be treated
    as constant memory. Therefore, there are several checks to be made
    before even running the algorithim.
    
    Additionally, the implemented algorithim must defend against cold-boot
    attacks, so the usage of memory that clears itself after being used
    is a must. This is especially important where the encryption key
    was stored. Therefore, a for loop is used to clear the allocated
    memory before the encryption function returns to the caller. 
    Also, the encryption program must defend against multiple other 
    kinds of attacks, most of which depend on finding the encryption 
    key sitting somewhere in memory, therefore the encryption keys 
    used in this program shouldn't look like normal keys at all, 
    to protect against such an attack.
    
    Now, when implementing the actual algorithim your function will
    require six parameters to be passed to it, if one is implementing 
    it in C/C++. Three parameters are pointers to hold the input, 
    output, and the key, the rest are to hold the amount of memory
    allocated to each pointer. When implementing the for loops, 
    a 64-bit counter will be necessary due to the possibly large
    amount of input data.

    \singlespacing
    \begin{lstlisting}[caption=C++ Encryption Source Code]
void encryptString(const uint8_t *KEY, uint16_t keyLen, uint8_t *instr,
                   uint64_t inlen, uint8_t *outstr, uint64_t outlen){
    assert(inlen <= outlen);
    assert(KEY != NULL);
    assert(instr != NULL);
    assert(outstr != NULL);
    uint8_t *KeyCopy = new uint8_t[keyLen];
    memcpy(KeyCopy, KEY, keyLen);
    for(uint64_t i = 0; i < inlen; i++){
        outstr[i] = instr[i] ^ KeyCopy[i % keyLen];
        outstr[i] = rotateRight(outstr[i], 1);
        if((i % keyLen) == 0){
            for(int j = 0; j < keyLen; j++){
                KeyCopy[j] = rotateRight(KeyCopy[j], 1);
            }
        }
    }
    // Secure Memory
    for(uint16_t i = 0; i < keyLen; i++){
        KeyCopy[i] = 0; // overwrite the key copy
    }
    delete[] KeyCopy;
}
    \end{lstlisting}

    

\doublespacing
\subsection{Decrypting a message}
    Decryption is just undoing what has been done to a message, now a
    good cipher will have many differint possibilities to defeat most
    cryptoanalysis methods. In other words, decryption is just inverse
    operations. However, this doesn't work so well if one is trying to
    break an encryption algorithim, as  a good algorithim would use a
    key and have hundreds of millions of possibilities. Thus, rendering
    a brute force approach impossible. As such to decrypt our message,
    we need 

    However, decryption is just an inversion of the encyption algorithim,
    as such we can implement it in revese as such, see the listing below.

    \begin{figure}[h]
    \begin{algorithmic}[1]
    \For{ $i=0; i < length(EncryptedMessage); i++$}
      \State $EncryptedMessage_i = rotateLeft(EncryptedMessage_i)$
      \State $DecryptedMessage_i = EncryptedMessage_i \oplus 
                                   Key_{i \bmod length(Key)}$
      \If{$(i \bmod length(Key)) \equiv 0$}
        \For{ $j=0;j < length(Key); j++$ }
          \State $Key_j = rotateRight(Key_j)$
        \EndFor
      \EndIf
    \EndFor
    \end{algorithmic}
    \caption{A reversal of an XOR rotation cipher}
    \end{figure}

    To undo the encryption algorithim, the key used to encrypt the message
    originally, and the exact operations used to encrypt the message must
    be done in reverse, however the operations performed on the key must
    be exactly the same operations originally if the encrypted message
    is being iterated forwards, in order to keep the key consistant 
    with the message.
    
    \begin{equation}
      \label{eq:XORinversion}
      (n \oplus k) \oplus k = n
    \end{equation}

    As the operations performed to decrypt a message using an XOR rotation
    cipher must be done in reverse from the encrpytion, meaning the first
    operation performed is bitwise rotate left. Then, the resulting value
    have an XOR operation performed on it with the key used for that index
    during encryption. As XOR is its own inverse operation as shown in
    equation \eqref{eq:XORinversion}, the XOR operation remains in the
    decryption algorithim.

    \subsubsection{Programming a Decryption Algorithim}
    Implementing a decryption algorithim is a bit more challenging than
    implementing an encryption algorithim. As some encrypted data can
    include some padding in the actual message, or in other words 
    garbage information. Recognizing the padding can be a major
    challenge, as some messages might have padding, while other 
    messages might not.

    \singlespacing
    \begin{lstlisting}[caption=C++ Decryption Code]
void decryptString(const uint8_t *KEY, uint16_t keyLen, uint8_t *instr,
                   uint64_t inlen, uint8_t *outstr, uint64_t outlen){
    assert(inlen <= outlen);
    assert(KEY != NULL);
    assert(instr != NULL);
    assert(outstr != NULL);
    uint8_t *KeyCopy = new uint8_t[keyLen];
    memcpy(KeyCopy, KEY, keyLen);
    for(uint64_t i = 0; i < inlen; i++){
        outstr[i] = rotateLeft(instr[i], 1);
        outstr[i] = outstr[i] ^ KeyCopy[i % keyLen];
        if((i % keyLen) == 0){
            for(int j = 0; j < keyLen; j++){
                KeyCopy[j] = rotateRight(KeyCopy[j], 1);
            }
        }
    }
    // Secure Memory
    for(uint16_t i = 0; i < keyLen; i++){
        KeyCopy[i] = 0; // overwrite the key copy
    }
    delete[] KeyCopy;
    //std::cerr << "Finished Decrypting String" << std::endl;
}
    \end{lstlisting}
    \doublespacing
 
\section{Stenography}
  Stenography is the practice on hiding a message. The goal is take
  a message and make sure that it gets to where it needs to get by
  making the message appear to not be a message. After all, if one
  does not know a message exists, they can not intercept it. Therefore,
  stenography should be an intergral part of the message sending process.
  
  \subsection{Image Stenography}
    As images are a very common feature in the modern internet browsing
    experience, making it not hard to see that hiding a message inside 
    of an image is a great idea. However, most images found in 
    modern web pages are compressed images, which can either be either
    lossless or lossy. Lossy compression is bad for stenography in 
    images as it causes the loss of data in the image, and can
    corrupt the message. As such lossless compression, which doesn't 
    change the byte values of each pixel, thus making it superior 
    for stenography.
    
    The easist way to think of images is as a multidimensional array.
    Which is pretty much exactly what images are. As all images are
    essentially a three dimensional array. With two dimensions being
    length and width, and another dimension being the color at that 
    point.

    \subsubsection{Hiding a Message}
      A message can be hidden in inside of any image. However it is best 
      to hide a message inside of a section of an image where the pixel
      values surrounding the message will hide the message. As the goal
      is to make the hidden message to look like it belongs in that
      picture, the colors must match up. As such an area must be found
      in the hidding place for the message. However, as it would be
      computationally expensive to search for it by computer, the
      messages values are averaged to find the average RGB values 
      instead, and the user specifies where to put the message 
      inside the image.
      
      Additionally, there are many patterns that the message can be
      written out as. A message can be written out as rows, columns,
      or a set of concentric circles. It doesn't really matter so long
      as the pixels are read back in the same order as they were
      written out.
          
    \subsubsection{Recovering a Message}    
      Recovering a message is not that complicated. As all that needs
      to be done is read out the pixel values in the exact order they
      were placed into the image. As such it can be recovered with a
      simple nested FOR loop. Then copying those values into an array
      of sorts. After that a transformation is ran to recover the
      original data.

    \subsubsection{Implementing Image Stenography}
      Implementing stenography with images is quite a challenge due to
      the varity of image formats available on the internet. Therefore,
      using a libray to abstract away file I/O away is neccessary.  As
      such, the CImg library provides a simple interface to do this
      \cite{cimg}. Additionally, it provides a real easy way to think of
      images as just multidimensional arrays, and it provides a simple
      class with extremely powerful set of methods to work with images.
      Thus, making the CImg library the best choice for implementing 
      stenography with images.

      
% Biblography
\bibliographystyle{plain}
\bibliography{Encryption.bib}
\end{document}
